{\bf Topology, Geometry and Morphology of the Dark Matter Web}

### 2018-07-24 Draft/Chapter6/concl.tex
final
final changes

@@

% \begin{itemize}
\subsection{Hierarchical structure of the dark matter web}
Hierarchical structure of the dark matter web: The first study of the multi-stream environment of dark matter haloes in cosmological N-body simulations in the $\Lambda$CDM cosmology reported in Chapter \ref{chapter3} (first shown in \cite{Ramachandra2015}) shows that at the resolution of the simulation i.e. without additional smoothing, the cosmic web represents a hierarchical structure: each halo is embedded in the filamentary framework of the web predominantly  at the filament crossings, and each filament is embedded in the wall like fabric of the web at the wall crossings.

Voids are uniquely defined by the local condition requiring to be a single-stream flow region. Within the dark matter web (regions with $n_{str} > 1$), the walls are regions where $n_{str} \cong 3$. The number of streams in the neighboring filaments is higher than in the neighboring walls. Finally, each halo or sub-halo is a local peak in the $n_{str}$ field. We demonstrated in Chapter \ref{chapter3} that the shells of streams around haloes are quite thin and the closest void region is typically within one and a half FOF radius from the center of the halo.

\hl{The multistream web represents a nested structure, consisting of layers with increasing number of streams.}

@@

Multistream view of the dark matter haloes: In our simulation reported in Chapter \ref{chapter3}, simple virial density criterion corresponds to a global threshold of $n_{str} \geq 90$ for DM haloes. This shows a reasonably good correspondence with structure finders. A more sophisticated halo finding formulation is done in Chapter \ref{chapter5} assuming that the virialized haloes have convex boundaries. Closed and convex regions of the multistream field are hence isolated by imposing a positivity condition on all three eigenvalues of the Hessian estimated on the smoothed multistream field. In a single-scale analysis of high multistream field resolution and low softening length, the halo substructures with local multistream maxima are isolated as individual halo sites. This halo finding technique is free of any heuristic parameters, and may be extended to delineate filaments, walls, voids, as well as halo sub-structures. Agreement of halo summary statistics (like the halo mass function) of the multistream haloes with traditional halo finders (Friends-of-Friends and Spherical Over-density based techniques) is shown \cite{Ramachandra2017b}. As opposed to several other halo finders, this geometrical approach accounts for non-spherical boundaries, does not find any halo candidates in voids, and minimizes the need for post-processing in the production of halo catalogs for dark matter simulations.

@@

Topological transitions in the cosmic web: Topological connections in the single-streaming voids and multistreaming filaments and walls reveal a cosmic web structure different from traditional mass density fields, as shown in Chapter \ref{chapter4}. Single streaming voids occupy over 90 per cent of the dark matter Universe. A single void structure not only percolates the multistream field in all the directions, but also spans over 99 per cent of all the single-streaming regions. Sub-grid analyses on scales smaller than simulation resolution reveal tiny pockets of voids that are isolated by membranes of the structure.

>>>>

% 2017 Paper_b abstract

% 2017 Paper_a
% Topological connections in the single-streaming voids and multistreaming filaments and walls reveal a cosmic web structure different from traditional mass density fields. A single void structure not only percolates the multistream field in all the directions, but also occupies over 99 per cent of all the single-streaming regions. Sub-grid analyses on scales smaller than simulation resolution reveal tiny pockets of voids that are isolated by membranes of the structure. For the multistreaming excursion sets, the percolating structure is significantly  thinner than the filaments in over-density excursion approach.  
% 
% Hessian eigenvalues of the multistream field are used as local geometrical indicators of dark matter structures. Single-streaming regions have most of the zero eigenvalues. Parameter-free conditions on the eigenvalues in the multistream region may be used to delineate primitive geometries with concavities corresponding to filaments, walls and haloes.

@@

% \begin{itemize}
\subsection{Hierarchical structure of the dark matter web}
% Hierarchical structure of the dark matter web: 
The first study of the multi-stream environment of dark matter haloes in cosmological N-body simulations in the $\Lambda$CDM cosmology reported in Chapter \ref{chapter3} (first shown in \cite{Ramachandra2015}) shows that at the resolution of the simulation i.e. without additional smoothing, the cosmic web represents a hierarchical structure: each halo is embedded in the filamentary framework of the web predominantly at the filament crossings, and each filament is embedded in the wall like fabric of the web at the wall crossings.

Voids are uniquely defined by the local condition requiring to be a single-stream flow region. This definition unambiguous and physically motivated, i.e., it implies that the voids are regions in the Universe where a gravitational collapse has not occurred. The single-streaming voids occupies around 90 per cent of total simulation volume and about 30 per cent of total mass. This results in over-density in the voids as $\delta_{v} \approx -0.7$, in agreement with several other definitions on voids.

On the other hand, such definitions for walls, filaments and haloes are empirical, with or without the multistream picture. Haloes, for instance, are often detected using proxies for dark matter particle proximity in position space. These definitions can be simple mass density thresholds, or based on geometry of the dark matter structure (spheroidal, tubular or flat) or topology (connected/isolated).

Our algorithm in \cite{Ramachandra2015} was motivated by the complex and diverse morphologies exhibited by the multi-stream field around a halo. We isolated various halo environments, and checked the variation of areal fraction of multistreams on concentric shells around the halo centers, with radii $r = 0.1 h^{-1} Mpc$ to $5 h^{-1} Mpc$ for each of these haloes. If this fraction scales as $f_{str}(r) \propto r^{-1}$ with respect to the radius $r$, the corresponding multistream region is geometrically identical to that of a wall or a pancake. Such scaling is seen for $n_{str} \cong 3$ regions. This result is completely in agreement of ZA, where the pancakes are the first collapsed regions formed. Our morphology based algorithm also identifies regions with three streams as pancakes.

Between multistream values $5-15$, the scaling gradually changes from $f_{str}(r) \propto r^{-1}$ to $f_{str}(r) \propto r^{-2}$. Particularly for our simulation, the variation of $n_{str} \cong 17$ is closest to that of filamentary geometry i.e., $r^{-2}$. As discussed in Chapter \ref{chapter4}, $n_{str} \cong 17$ is also the value at which rapid percolation transition from isolated islands to connected filaments occurs in the multistream field.

Finally, at multistream thresholds $n_{str} > 17$, the variation $f_{str}(r)$ departs from $r^{-2}$ towards that of knots. The generally accepted mass density threshold for compact haloes, the virial over-density of $\rho_{vir} = 200$ corresponded to $n_{str} \cong 90 $ in our simulation. We also demonstrated in Chapter \ref{chapter3} that the shells of streams around haloes are quite thin and the closest void region is typically within one and a half FOF radius from the center of the halo.

% Within the dark matter web (regions with $n_{str} > 1$), the walls are regions where $n_{str} \cong 3$. 
Qualitatively, the results are in good agreement with the hierarchical picture of the large scale structures: The multistream web represents a nested structure, consisting of layers with increasing number of streams. The number of streams in the neighboring filaments is higher than in the neighboring walls. However, some of the quantitative results tabulated above (and mentioned in Chapter \ref{chapter3} in detail) may vary depending on the simulation parameters and cosmological models. With the exception of void definition as regions with $n_{str} = 1$, one should be cautious about using the same multistream threshold for walls, filaments and haloes. Instead, the heuristic multistream cut-offs should be verfied or re-calculated for the specific cosmological simulation.

@@

% Multistream view of the dark matter haloes: 
In our simulation reported in Chapter \ref{chapter3}, simple virial density criterion corresponds to a global threshold of $n_{str} \geq 90$ for DM haloes. This shows a reasonably good correspondence with structure finders. A more sophisticated halo finding formulation is done in Chapter \ref{chapter5} assuming that the virialized haloes have convex boundaries. Closed and convex regions of the multistream field are hence isolated by imposing a positivity condition on all three eigenvalues of the Hessian estimated on the smoothed multistream field. In a single-scale analysis of high multistream field resolution and low softening length, the halo substructures with local multistream maxima are isolated as individual halo sites. This halo finding technique is free of any heuristic parameters, and may be extended to delineate filaments, walls, voids, as well as halo sub-structures. Agreement of halo summary statistics (like the halo mass function) of the multistream haloes with traditional halo finders (Friends-of-Friends and Spherical Over-density based techniques) is shown \cite{Ramachandra2017b}. As opposed to several other halo finders, this geometrical approach accounts for non-spherical boundaries, does not find any halo candidates in voids, and minimizes the need for post-processing in the production of halo catalogs for dark matter simulations.

@@

% Topological transitions in the cosmic web: 
Topological connections in the single-streaming voids and multistreaming filaments and walls reveal a cosmic web structure different from traditional mass density fields, as shown in Chapter \ref{chapter4}. Single streaming voids occupy over 90 per cent of the dark matter Universe. A single void structure not only percolates the multistream field in all the directions, but also spans over 99 per cent of all the single-streaming regions. Sub-grid analyses on scales smaller than simulation resolution reveal tiny pockets of voids that are isolated by membranes of the structure.
